A quantum parallel Markov chain Monte Carlo

### 2021-11-05 main.tex
qunatum state as probability

\newcommand{\ttheta}{\boldsymbol{\theta}}
\newcommand{\dd}{\mbox{d}}

@@

Quantum computers perform operations on unit-length vectors of complex data called \emph{qubits} or quantum bits.  One may write any qubit $\ket{\psi}$ as a linear combination of the \emph{computational basis} states $\ket{0}$ and $\ket{1}$.  In symbols,
\begin{align*}
\ket{\psi} = \alpha \ket{0} + \beta \ket{1} \quad \mbox{for} \quad \alpha, \, \beta \in \mathbb{C} \quad \mbox{and} \quad  |\alpha|^2 + |\beta|^2 = 1\, .
\end{align*}
This formula makes use of \emph{Dirac} or \emph{bra-ket} notation and thinly obscures ideas that are commonplace in the realm of applied statistics: we may make the unit-vector specification of $\ket{\psi}$ clear by writing
\begin{align*}
\ket{0}=\begin{bmatrix}
1 \\ 0
\end{bmatrix} \, ,  \quad \ket{1}=\begin{bmatrix}
0 \\ 1
\end{bmatrix} \quad \mbox{or} \quad \ket{\psi} = \alpha \begin{bmatrix}
1 \\ 0
\end{bmatrix} + \beta  \begin{bmatrix}
0 \\ 1
\end{bmatrix}\, .
\end{align*} 
As one might expect, the full machinery of linear algebra is also available. The conjugate transpose of $\ket{\psi}$ is $\bra{\psi}$. The inner product of $\ket{\psi}$ and $\ket{\phi}$ is $\braket{\phi|\psi}$. The outer product is $\ket{\psi}\bra{\phi}$.

@@

@@

\section{Quantum optimization}

\section{Algorithms}

>>>>

\newcommand{\ttheta}{\boldsymbol{\theta}}
\newcommand{\dd}{\mbox{d}}
\newcommand{\ppsi}{\boldsymbol{\psi}}
\newcommand{\U}{\mathbf{U}}
\newcommand{\I}{\mathbf{I}}
\renewcommand{\H}{\mathbf{H}}

@@

Quantum computers perform operations on unit-length vectors of complex data called \emph{qubits} or quantum bits.  One may write any qubit $\ppsi$ as a linear combination of the \emph{computational basis states} $\ket{0}$ and $\ket{1}$.  In symbols,
\begin{align*}
\ket{\psi} = \alpha \ket{0} + \beta \ket{1} \quad \mbox{for} \quad \alpha, \, \beta \in \mathbb{C} \quad \mbox{and} \quad  |\alpha|^2 + |\beta|^2 = 1\, .
\end{align*}
This formula makes use of \emph{Dirac} or \emph{bra-ket} notation and thinly obscures ideas that are commonplace in the realm of applied statistics: we may make the unit-vector specification of $\ket{\psi}$ clear by writing
\begin{align*}
\ket{0}=\begin{bmatrix}
1 \\ 0
\end{bmatrix} \, ,  \quad \ket{1}=\begin{bmatrix}
0 \\ 1
\end{bmatrix} \quad \mbox{or} \quad \ket{\psi} = \alpha \begin{bmatrix}
1 \\ 0
\end{bmatrix} + \beta  \begin{bmatrix}
0 \\ 1
\end{bmatrix}\, .
\end{align*} 
As one might expect, the full machinery of linear algebra is also available within this notation. The conjugate transpose of $\ket{\psi}$ is $\bra{\psi}$. The inner product of $\ket{\psi}$ and $\ket{\phi}$ is $\braket{\phi|\psi}$. The outer product is $\ket{\psi}\bra{\phi}$. Naturally, we can write $\ppsi$ as a linear combination of any other such basis elements. Consider instead the vectors
\begin{align*}
\ket{+} = \frac{1}{\sqrt{2}} \ket{0} + \frac{1}{\sqrt{2}} \ket{1} \quad \mbox{and} \quad \ket{-} = \frac{1}{\sqrt{2}} \ket{0} - \frac{1}{\sqrt{2}} \ket{1} \, .
\end{align*}
A little algebra shows that $\ket{+}$ and $\ket{-}$ are indeed unit-length and orthogonal to each other. A little more algebra reveals that, with respect to this basis, $\ppsi$ has coefficients $\alpha'$ and $\beta'$ given by $(\alpha + \beta)/\sqrt{2}$ and $(\alpha - \beta)/\sqrt{2}$, respectively. A last bit of algebra shows that this representation is consistent with $\ppsi$'s unit length.

But linear algebra is not the only aspect of quantum computing that should come easily to applied statisticians. In addition to thinking of $\ppsi$ as a vector, it is also useful to think of $\ppsi$ as a (discrete) probability distribution over the computational basis states $\ket{0}$ and $\ket{1}$. The constraints on coefficients $\alpha$ and $\beta$ mean that we can think of $|\alpha|^2$ and $|\beta|^2$ as probabilities that $\ppsi$ is in state $\ket{0}$ or $\ket{1}$, respectively.  Accordingly, $\ket{+}$ and $\ket{-}$ encode uniform distributions over the computational basis states.  In the parlance of quantum mechanics, $\alpha$, $\beta$ and $\pm1/\sqrt{2}$ are all \emph{probability amplitudes}, and $\ppsi$, $\ket{+}$ and $\ket{-}$ are all \emph{superpositions} of the computational basis states. Quantum \emph{measurement} of $\ppsi$ results in $\ket{0}$ with probability $|\alpha|^2$, but---in the following---we can safely think of this physical procedure as drawing a single discrete sample from $\ppsi$'s implied probability distribution.

==============================

Quantum logic gates perform linear operations on qubits and take the form of unitary matrices $\U$ satisfying $\U^\dagger \U=\I$ for $\U^\dagger$ the conjugate transpose of $\U$. One terrifically important quantum gate is the \emph{Hadamard gate}
\begin{align}
\H = \frac{1}{\sqrt{2}} \begin{bmatrix}
1 & 1 \\
1 & -1 \end{bmatrix} \, .
\end{align}
We invite the Reader to check for themselves that $\H$ is indeed unitary and that its action maps $\ket{0}$ to $\ket{+}$ and $\ket{1}$ to $\ket{-}$.

=============================

@@

=============================

@@
